The Galaxy Clustering Crisis in Abundance Matching

### 2017-05-07 paper/paper.tex
updates

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]
{\parbox[t]{\textwidth}{
Duncan Campbell$^{1}$\thanks{E-mail: duncan.campbell@yale.edu},
Frank C. van den Bosch$^{1}$,
Nikhil Padmanabhan$^{2, 3}$, \\ 
Andrew Zentner$^{4}$,
Johannes Lange$^{1}$,
Yao-Yuan Mao$^{4}$, and
Fangzhou Jiang$^{5}$
} \\
\vspace*{3pt} \\
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
$^{3}$Yale Center for Astronomy and Astrophysics, Yale University, New Haven, CT 06520, USA\\
$^{4}$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260\\
$^{5}$Racah Institute of Physics, The Hebrew University, Jerusalem 91904, Israel
}

@@

In this section, we examine the origin of the success of $V_{\rm peak}$ based SHAM models.  By comparing the differences in $V_{\rm peak}$ and $M_{\rm peak}$ based models, we identify the culprit(s) in the failure of the mass-based models to match observational clustering measurements.  We also discuss some reasons to prefer mass-based models over $V_{\rm peak}$ based models.

\subsection{Satellite Fraction}

Where clustering is under-predicted, the problem is most severe at small scales $(<1~h^{-1}{\rm Mpc})$.  This is a strong indication that the culprit is satellite galaxies, or more precisely, a lack of satellite galaxies in the mass-based SHAM mocks.  With this in mind, we examine the satellite fraction, $f_{\rm sat}$, in each mock in Fig. \ref{fig:f_sat_mstar}.  The RV model has a larger satellite fraction, $\sim 27\%$, relative to the other models which are below $\sim 22\%$.  Comparing the clustering signals in Fig. \ref{fig:wp_comparison} with the satellite fractions in Fig. \ref{fig:f_sat_mstar} shows that there is a nearly one-to-one correspondence between the clustering strength on small scales and $f_{\rm sat}$ of each model.  Furthermore, the satellite fractions between the models do not diverge until approximately $5 \times 10^{10}h^{-2}M_{\odot}$, above which the models largely agree with the clustering results.  Considering this, it may be more appropriate to restate the under-prediction of galaxy clustering in these models as an underproduction of satellite galaxies.

\subsection{Why does $V_{\rm peak}$ SHAM have increased $f_{\rm sat}$?}

As can be seen in Fig. \ref{fig:mpeak_vpeak}, $V_{\rm peak}$ and $M_{\rm peak}$ are highly correlated properties with small scatter, $\sigma_{\log(V_{\rm peak})}\sim 0.05$ at fixed mass for host-haloes.  Additionally, sub-haloes have on average larger $V_{\rm peak}$ than host-haloes of equal $M_{\rm peak}$.  The reasoning behind this difference between host-haloes and sub-haloes is straight forward.  For an NFW profile, $V_{\rm max}$ is an increasing function of the halo mass and concentration, $V_{\rm max} = f(m_{\rm vir}, c_{\rm vir})$ \citep[see eq. 6 in][]{Bosch:2014cu}. Furthermore, it is a robust prediction of $\Lambda$CDM dark matter simulations that halo concentration correlates with halo formation time, $\langle z_f \rangle = f(m_{\rm vir}, c_{\rm vir})$ \citep[e.g.][]{Wechsler:2002kh}.  Finally, because $M_{\rm peak}$ occurs before accretion for sub-haloes and near z=0 for host-haloes, sub-haloes generally have earlier formation times at fixed $M_{\rm peak}$.

Given this, and using the concentration-mass-redshift relation from \citet{Maccio:2009bs}.  We devise a description for the median $M_{\rm peak}-V_{\rm peak}$ relation given by:
%
\begin{linenomath}
\begin{equation}
\label{eq:vpeak}
\langle V_{\rm peak}\rangle (M_{\rm peak}, z) = 1.1\times V_{\rm max}(M_{\rm peak}, z)
\end{equation}
\end{linenomath}
%
where we include the factor of $1.1$ to account for the fact that the average peak maximum circular velocity is $\sim 10\%$ higher than $V_{\rm max}$ \citep{Behroozi:2014kk}.  We show this relation at $z=0$ (solid line) and at $z=1.0$ (dashed line) in Fig. \ref{fig:mpeak_vpeak}.

It is the correlated scatter in $V_{\rm peak}$ at fixed $M_{\rm peak}$ which is responsible for the difference in the satellite fractions between the RV and RM models.  By populating (sub-)haloes by their rank on $V_{\rm peak}$, more sub-haloes will be populated than would have been had $M_{\rm peak}$ instead been utilized for any given stellar mass threshold sample.  This same reasoning carries over to the other mass-based models.

@@

Here we compare the RV model to the other models by examining the relation between stellar mass and halo mass that arises in each model, in the RV model by correlation between $V_{\rm peak}$ and $M_{\rm peak}$, and in the other models by design.  In the top panel of Fig. \ref{fig:SMHM_comparison}, we examine the relation between stellar mass and $M_{\rm peak}$ for individual galaxies in the RV mock.  Beyond the ubiquitous shape of the SMHM relation, the most obvious additional trend is that the RV model predicts that sub-haloes ($z_{\rm acc}>0$) host more massive galaxies than host-haloes of equivalent $M_{\rm peak}$.  This is precisely the expectation given that sub-haloes generally have larger values of $V_{\rm peak}$ than host-haloes as shown in the previous section.

We show the same relations for the M13, B13, and Y12 models in the lower panels of Fig. \ref{fig:SMHM_comparison}, but instead of showing it for individual galaxies, we show the mean relation as a function $z_{\rm acc}$ for simplicity and ease of interpretation.  Broadly, to the degree that there is a difference in the relation for centrals and satellites, each of these models predicts the qualitative opposite trend as the RV model; sub-haloes host {\em less} massive galaxies than host-haloes of equivalent $M_{\rm peak}$ at masses less than $\sim 10^{12.5} ~ h^{-1} M_{\odot}$.  In the M13 and B13 models, this comes about because the SMHM relation for satellites as a function of $z_{\rm acc}$ is a fossil record of the relation for centrals at higher redshift since it is assumed that satellites cease to grow appreciably after they are accreted.

On the other hand, the Y12 model predicts almost no difference in the stellar mass between satellites and centrals at fixed halo mass.  This is because in the Y12 model satellites are allowed to grow (or lose stellar mass) after accretion.  We show the relation at $z=0.5$ and $1.0$ for {\em centrals} as a dotted line in the Y12 panels in Fig. \ref{fig:SMHM_comparison}.  It is very interesting that the parameter constraints found by Y12 indicate that this happens post accretion evolution happens very efficiently, with satellites achieving a final stellar mass $> 90\%$ that of centrals at $z=0$.  This phenomenology also explains why Y12 predicts a slightly larger satellite fraction than the M13 and B13 models.  The sub-halo fraction increases as $M_{\rm peak}$ decreases (see Fig. \ref{fig:f_sub}), resulting in more satellites as lower $M_{\rm peak}$ sub-haloes are populated.  The same is true for the RM model compared to the M13 and B13 models.  However, this effect in the Y12 and RM models alleviates the clustering decrement only slightly relative to the M13 and B13 models.  In the absence of post-accretion evolution of satellites in Y12(c parameter set to 0), the model looks similar to M13 and B13, where satellites are less massive than centrals at z=0.0.  However, it is telling that Y12 strongly prefers large values for $c$ in the analysis performed in \citet[][]{Yang:2012ew}.

@@

Given that the RV and mass-based models make different assumptions for the SMHM relation, especially regarding satellites, it is not be surprising that each model predicts different galaxy clustering signals and satellite abundances.   Considering the RV model's success in fitting clustering observations and the mass-based models failure (see \S \ref{sec:clustering}), it appears that the RV model should be favoured as the more physical.  However, in this section we show that the RV model implies galaxy growth histories that are incompatible with both the other models and observations.

% Up to this point, one might be inclined to favour the RV model over the mass-based SHAM models given its success in fitting galaxy clustering observations.

In the previous section we showed that, in the RV model, the stellar masses of satellites are on average larger than centrals at fixed $M_{\rm peak}$.  This finding implies that the SMHM relation in the RV model decreases over time (evolves towards the right in Fig. \ref{fig:SMHM_comparison}), while it increases in the other models (evolves from right to left).  This conclusion rests on two implicit assumptions of the RV model: the stellar mass of satellites does not evolve significantly after $z_{\rm acc}$, and the $M_* - V_{\rm peak}$ (SMVP) relation itself does not evolve.  A non-evolving SMVP implies that galaxies grow along the SMVP, i.e. galaxy growth is tied to the growth of $V_{\rm peak}$.  In this section we examine the latter assumption, and we leave an examination of post-accretion satellite evolution to \S \ref{sec:sat_growth}.

@@

Fig. \ref{fig:mstar_of_z_rm} shows that, in the RV model, galaxy growth is largely self-similar--at any given redshift, galaxies of every mass are growing at similar rates.  Conversely, in the M13 model, galaxy growth is much more dependent on halo mass--high mass galaxies form early, growing slowly at low redshift, while low mass galaxies form late.  Slow late time growth of massive galaxies is necessary to reproduce the observed prevalence of quiescent galaxies as mass increases \citep[e.g.][]{Wetzel:2012lk}.  It is the same story when comparing RV to any of the other mass based models.  In Fig. \ref{fig:mstar_zform}, we show the the median redshift galaxies found it different mass haloes at $z=0$ reach a fraction, $f$, of their $z=0$ stellar mass.  Mass-based SHAM models generically predict much later growth compared to RV.  The one exception to this rule is the Y12 model, where at late times, massive galaxies lose stellar mass.  As a result, galaxies found in massive haloes reached $90\%$ of their $z=0$ mass relatively early.  Since we are not concerned with the evolution of massive galaxies, we do not investigate this aspect of Y12 any further in this paper.  Interestingly, even the RM model, where the SMHM relation does not evolve, results in galaxy growth much closer to the evolving models than with RV.  This is a consequence of the fact that haloes form their potential wells early, primarily adding mass in the outskirts at late times \citep{Bosch:2014cu}.  As a result, any model for galaxy formation based on $V_{\rm peak}$ will have to decouple galaxy growth from the growth of $V_{\rm peak}$ more than models based on $M_{\rm peak}$.

@@

Given that the M13, Y12, and B13 models were fit to the stellar mass function at various redshifts and reproduce the cosmic star-formation density in the Universe (among other observables), it is difficult to imagine how to make a model like RV consistent with these same observational constraints.  While not explored here, it is likely that a more complicated $V_{\rm peak}$ based SHAM model could produce realistic galaxy growth histories, but this would require evolution in the SMVP relation.  Any evolution will require the SMVP relation be lower at high redshifts (as is the case for the SMHM relation in M13, Y12, and B13), reducing the stellar mass assigned to satellite galaxies in typical implementations.  This sort of modification would result in a decrease in the galaxy clustering signal on small scales.

Given this, it is not clear that there is a reason to prefer a $V_{\rm peak}$ based SHAM model over one based on $M_{\rm peak}$.  Apart from arguments related to fitting $w_p(r_p)$ and $\phi(M_*,z)$, there are other difficulties that must be overcome if $V_{\rm peak}$ is to be argued as more fundamental driving the stellar mass mapping into haloes.  For any given (sub-)halo, $V_{\rm peak}$ is generally set during major mergers \citep[1:5 or larger, ][]{Behroozi:2014bg}, and if stellar mass is tightly correlated with $V_{\rm peak}$, this would imply bursty star-formation closely tied to major mergers which is not favoured by observations \citep[e.g. star formation histories][]{Diemer:2017gj}.  On the other-hand, $M_{\rm peak}$ remains a theoretically attractive as a quantity that should be tightly coupled to stellar mass.  Peak halo mass is a good indicator of the amount of baryons that have been available to a galaxy to grow over its history {\bf [citations needed]}.

In our comparison between $V_{\rm peak}$-based SHAM and $V_{\rm peak}$-based SHAM it is clear that there is a tension between fitting galaxy clustering on small scales while simultaneously reproducing the build up of stellar mass in the Universe.  This tension presents a ``clustering crisis'' for SHAM.  Our exploration in the preceding sections has suggests multiple avenues SHAM models could be altered in order to alleviate this crisis.  We have fingered too few satellites as the culprit for the clustering deficiency in mass based SHAM models.  One solution is to simply increase the number of satellites in the mass-based models as many models have found to be necessary.  In \S \ref{sec:orphans}, we examine the plausibility of missing sub-haloes, and thus satellites, in our implementation of the mass-based SHAM models.  A significant population of missing sub-haloes would be an indication that so called ``oprhan'' galaxies play an important role in solving the clustering crisis.  Apart from orphans, we also consider two other physically motivated modifications to increase the satellite fraction--post-accretion satellite growth in \S \ref{sec:sat_growth} and assembly bias effects in \S \ref{sec:assem_bias}.

@@

One possible solution to the lack of small scale clustering signal in SHAM models is to include a population of ``orphan'' galaxies.  Given the finite mass resolution of dark matter simulations, it is reasonable to expect that some sub-haloes are artificially disrupted or otherwise missing from the halo catalogues at z=0 \citep{Guo:2013fm}.  Alternatively, sub-halo finders may fail in identifying sub-haloes when the density contrast is low \citep{Wetzel:2010ei, Muldrew:2011gr, Knebe:2011jc, Onions:2012iv, Knebe:2013bp, vandenBosch:2016ky} as is the case in the central regions of host haloes.  If this is in fact occurring, then it is appropriate to include a population of ``orphan'' galaxies, galaxies that have no identifiable sub-halo within a simulation but should rightfully be included if the simulation or sub-halo finder had been more successful.

After sub-haloes are accreted onto a more massive host-halo, mass is tidally stripped as the sub-halo orbits within the potential of its host, resulting in a ratio between the $z=0$ mass and the mass at accretion that is generally less than unity:
%
\begin{linenomath}
\begin{equation}
f_m = m_{\rm sub}/m_{\rm acc}
\end{equation}
\end{linenomath}
%
where the mass of a sub-halo, $m_{\rm sub}$, is the instantaneous mass that is bound to the sub-halo, and $m_{\rm acc}$ is the viral mass at the time of accretion.  Eventually, sub-haloes may simply be stripped below the mass resolution of the simulation, $f_m \times m_{\rm acc} \sim m_p$.  If it is common for sub-haloes to survive to this point, this will result in a need for orphans.  Alternatively, if sub-haloes are not tracked accurately, mass resolution is not sufficiently high, or the force resolution is not sufficient, some sub-haloes may become disrupted prematurely, meaning fewer sub-haloes will be available to host satellite galaxies when applying an abundance matching scheme.

@@

We calculate the sub-halo completeness in the simulation as the ratio of the empirical $f_{\rm sub}$ and the fit using eq. \ref{eq:fsub_fit} as shwon in the bottom panel of Fig. \ref{fig:f_sub}.  We then model the completeness, $C_{\rm sub}$, as a function of halo mass as:
%
\begin{linenomath}
\begin{equation}
\label{eq:sub_completeness}
C_{\rm sub}(M_{\rm peak}) = \frac{C_0}{1.0 + \left(\frac{M_0}{M_{\rm peak}}\right)^{\gamma}}
\end{equation}
\end{linenomath}
%
We find $\log(M_0/h^{-1}M_{\odot})=9.980\pm0.003$ and $\gamma=1.27 \pm 0.02$ provide a good fit, as shown by the red line in the bottom panel in Fig. \ref{fig:f_sub}.  For now, we fix $C_0=1.0$, and we examine the possibility of $C_0<1.0$ at the end of this section.  The implicit assumption here is that when sub-haloes are well resolved, as is the case for massive (sub-haloes) at the time of infall, there should be no missing sub-structure.  Our estimation of the completeness is broadly consistent with the 1000 particle threshold found by \citet{Guo:2013fm} using the Millennium simulation suite.  Furthermore, we find that $C_{\rm sub}(M_{\rm peak})$ is very nearly constant with redshift ($z=0-4$).

@@

Given the success of our modifications to the RM model to introduce assembly bias to $M_*$, we now consider modifications to the evolving mass-based modes.  As discussed, each of the evolving models makes the assumption that satellite galaxies at the time of accretion have the same mass as central galaxies in haloes with equal $M_{\rm peak}$.  In these models, correlating $M_*$ with $z_{\rm f}$ will increase the mass of satellite galaxies if haloes which become sub-haloes form significantly earlier than haloes which remain host haloes.  This differs from the model for post-accretion satellite growth discussed in the previous section as satellites will already be more massive than centrals in similar haloes at the time of accretion.

>>>>

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]
{\parbox[t]{\textwidth}{
Duncan Campbell$^{1}$\thanks{E-mail: duncan.campbell@yale.edu},
Frank C. van den Bosch$^{1}$,
Nikhil Padmanabhan$^{2, 3}$,\\ 
Johannes Lange$^{1}$,
Yao-Yuan Mao$^{4}$, 
Fangzhou Jiang$^{5}$,
Antonio Villarreal$^{4}$,\\
and Andrew Zentner$^{4}$
} \\
\vspace*{3pt} \\
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
$^{3}$Yale Center for Astronomy and Astrophysics, Yale University, New Haven, CT 06520, USA\\
$^{4}$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260\\
$^{5}$Racah Institute of Physics, The Hebrew University, Jerusalem 91904, Israel
}

@@

In this section, we examine the origin of the success of $V_{\rm peak}$ based SHAM models.  By comparing the differences in $V_{\rm peak}$ and $M_{\rm peak}$ based models, we identify the culprit(s) in the failure of the mass-based models to match observational clustering measurements.  We conclude the section by pointing out some deficiencies in $V_{\rm peak}$ SHAM.

\subsection{The Satellite Fraction}

In the mass-based models the clustering signal is most severely under-predicted at small scales $(<1~h^{-1}{\rm Mpc})$.  This is a strong indication that the culprit is satellite galaxies--or more precisely, a lack of satellite galaxies in the mass-based SHAM models.  With this in mind, we examine the satellite fraction, $f_{\rm sat}$, in each model in Fig. \ref{fig:f_sat_mstar}.  The RV model results in the largest number of satellites, approaching $\sim30\%$ at $10^{9.5}~h^{-2}M_{\odot}$, relative to the other models which remain below  $\sim24\%$.  Comparing the clustering signals in Fig. \ref{fig:wp_comparison} with $f_{\rm sat}$ in Fig. \ref{fig:f_sat_mstar} shows that there is a nearly one-to-one correspondence between the clustering strength on small scales and $f_{\rm sat}$ in each model.  Furthermore, $f_{\rm sat}$ between the models does not diverge until approximately $5 \times 10^{10}h^{-2}M_{\odot}$, above which the models largely agree with the observed clustering signal.  Considering this, it may be more appropriate to restate the under-prediction of galaxy clustering in these models as an under-production of satellite galaxies.

\subsection{Why is $f_{\rm sat}$ larger in $V_{\rm peak}$ SHAM?}

As can be seen in Fig. \ref{fig:mpeak_vpeak}, $V_{\rm peak}$ and $M_{\rm peak}$ are highly correlated properties of haloes with small scatter, $\sigma_{\log(V_{\rm peak})}\sim 0.05$, at fixed $M_{\rm peak}$ for host-haloes.  Additionally, sub-haloes have larger $V_{\rm peak}$ on average than host-haloes of equal $M_{\rm peak}$.  This difference between host-haloes and sub-haloes is a result of the correlation between halo formation history and concentration.

Assuming an NFW profile, $V_{\rm max}$ is an increasing function of the halo mass and concentration, $c$:
%
\begin{linenomath}
\begin{equation}
V_{\rm max} = 0.465V_{\rm vir}\sqrt{\frac{c}{f(c)}}
\end{equation}
\end{linenomath}
%
where,
\begin{linenomath}
\begin{align}
V_{\rm vir} = ~&159.43~{\rm km/s}\left(\frac{M_{\rm vir}}{10^{12}h^{-1}M_{\odot}}\right)^{1/3}\left[\frac{H(z)}{H_0}\right]^{1/3} \\ \nonumber
 &\times \left[\frac{\Delta_{\rm vir}(z)}{178\Omega^{-1}_m}\right]^{1/6}
\end{align}
\end{linenomath}
and,
%
\begin{linenomath}
\begin{equation}
f(x) = \ln(1+x) - \frac{x}{1+x}
\end{equation}
\end{linenomath}
Furthermore, it is a robust prediction of simulations of $\Lambda$CDM structure formation that halo concentration correlates with halo formation time, $\langle z_f \rangle = g(M_{\rm vir}, c)$ \citep[e.g.][]{Wechsler:2002kh}.  Finally, because $M_{\rm peak}$ occurs before accretion for sub-haloes and near z=0 for host-haloes, sub-haloes generally have earlier formation times at fixed $M_{\rm peak}$.

Using the concentration-mass-redshift relation from \citet{Maccio:2009bs}.  We formulate a description for the median $M_{\rm peak}-V_{\rm peak}$ relation given by:
%
\begin{linenomath}
\begin{equation}
\label{eq:vpeak}
\langle V_{\rm peak} | M_{\rm peak}\rangle_{\rm med}(z) = 1.1\times \langle V_{\rm max}|M_{\rm peak}\rangle_{\rm med}(z)
\end{equation}
\end{linenomath}
%
The factor of $1.1$ accounts for the fact that the average peak maximum circular velocity is $\sim 10\%$ higher than $V_{\rm max}$ \citep{Behroozi:2014kk}.  We show this relation at $z=0$ (solid line) and at $z=1.0$ (dashed line) in Fig. \ref{fig:mpeak_vpeak}.  At fixed $M_{\rm peak}$, $\langle V_{\rm peak}\rangle_{\rm med}$ is larger at higher redshifts.

It is the correlated scatter in $V_{\rm peak}$ at fixed $M_{\rm peak}$ at $z=0$ which is responsible for the difference in satellite fractions between the RV and RM models.  By populating (sub-)haloes by their rank on $V_{\rm peak}$, more sub-haloes will be populated than would have been had $M_{\rm peak}$ instead been utilized for any given stellar mass threshold.  This same reasoning carries over to the other mass-based models.

@@

Here we compare the RV model to the other models by examining the relation between stellar mass and halo mass that arises in each model, in the RV model by correlation between $V_{\rm peak}$ and $M_{\rm peak}$, and in the other models by design.  In the top panel of Fig. \ref{fig:SMHM_comparison}, we examine the relation between $M_*$ and $M_{\rm peak}$ in the RV model.  The median stellar mass mass of central galaxies is shown as points with error bars indicating the log-normal scatter at fixed halo mass.  The colour-coded points show the stellar mass of satellites.  Beyond the ubiquitous broken power-law shape of the SMHM relation, sub-haloes ($z_{\rm acc}>0$) host more massive galaxies than host-haloes of equivalent $M_{\rm peak}$, with earlier accreted satellites having the largest masses.  This is precisely the expectation for satellites given the correlation between $V_{\rm peak}$ and formation time shown in the previous section.

We show the SMHM relation for the M13, B13, and Y12 models in the lower panels of Fig. \ref{fig:SMHM_comparison}; however, instead of showing individual satellite galaxies, we show the mean relation as a function $z_{\rm acc}$ for simplicity and ease of interpretation.  To the degree that satellites follow a different relation than centrals, the mass-based models predict the qualitative opposite trend as the RV model; sub-haloes host {\em less} massive galaxies than host-haloes of equivalent $M_{\rm peak}$ (at least for haloes less massive than $\sim 10^{12.5} ~ h^{-1} M_{\odot}$).  In the M13 and B13 models, this comes about because the SMHM relation for satellites as a function of $z_{\rm acc}$ serves as a fossil record of the SMHM relation for centrals at higher redshift, and \citet{Moster:2013ab} and \citet{Behroozi:2013fg} find that the SMHM relation must evolve such that $M_*$ at fixed $M_{\rm peak}$ increases.  This holds because it is assumed that satellites cease to grow after being accreted.

On the other hand, in the Y12 model there is no difference in $\langle M_*\rangle_{\rm med}$ between satellites and centrals at fixed $M_{\rm peak}$.  The Y12 model is different from M13 and B13 in that satellites are allowed to grow (or lose stellar mass) after $z_{\rm acc}$.  Therefore, all of the lines fall ontop of eachother in the bottom panel of Fig. \ref{fig:SMHM_comparison}.  We also show the SMHM relation for centrals at $z=0.5$ and $1.0$ in the Y12 model as a dotted lines for comparison.  \citet{Yang:2012ew} find that in the Y12 model post-accretion accretion evolution happens very efficiently--satellites achieve a final stellar mass $> 90\%$ that of centrals by $z=0$.  For simplicity, in this paper we have assumed that satellites obtain the same stellar mass as centrals ($c$ parameter set to 1.0 in Y12), consistent with the uncertainty found by \citet{Yang:2012ew}.

This model for satellite growth is why the Y12 model predicts a slightly larger $f_{\rm sat}$ than the M13 and B13 models as shown in Fig \ref{fig:f_sat_mstar}.  Because the sub-halo fraction increases as $M_{\rm peak}$ decreases (see \S \ref{sec:orphans} and Fig. \ref{fig:f_sub}), any model which boosts $M_*$ in sub-haloes, will result in a larger fraction of satellites for any given stellar mass threshold.  The same reasoning holds for the RM model, relative to M13 and B13, where centrals and satellites always follow the same SMHM relation.  However, this effect in the Y12 and RM models alleviates the clustering signal decrement only slightly compared to M13 and B13.  In the absence of post-accretion evolution in $M_*$ of satellites in Y12 ($c$ parameter set to 0), the Y12 model appears similar to the M13 and B13 models where satellites are less massive than centrals at z=0.0 (below $\sim 10^{12.5} ~ h^{-1} M_{\odot}$).  It is interesting that in the analysis performed in \citet[][]{Yang:2012ew}, they find $c \sim 1$ is strongly preferred over $c \sim 0$.  We examine post-accretion evolution of satellites in more detail in \S \ref{sec:sat_growth}.

@@

Given that the RV and mass-based models make different assumptions for the SMHM relation, especially regarding satellites, it is not surprising that each model predicts different galaxy clustering signals and satellite abundances.   Considering the RV model's success in fitting clustering observations and the mass-based models failure (see \S \ref{sec:clustering}), it appears that the RV model should be favoured as the more physical.  However, in this section we show that the RV model implies galaxy growth histories that are incompatible with both the other models and observations.

In the previous section we showed that, in the RV model, the stellar masses of satellites are on average larger than centrals at fixed $M_{\rm peak}$.  This finding implies that the SMHM relation in the RV model evolves in the opposite direction as the mass-based models.  This conclusion rests on two implicit assumptions of the RV model: the stellar mass of satellites does not evolve significantly after $z_{\rm acc}$, and the $M_* - V_{\rm peak}$ (SMVP) relation itself does not evolve.  A non-evolving SMVP implies that galaxies grow along the SMVP, i.e. galaxy growth is tied to the growth of $V_{\rm peak}$.  In this section we examine the latter assumption, and we leave an examination of post-accretion satellite evolution to \S \ref{sec:sat_growth}.

@@

Fig. \ref{fig:mstar_of_z_rm} shows that, in the RV model, galaxy growth is largely self-similar--at any given redshift, galaxies of every mass are growing at similar rates.  Conversely, in the M13 model, galaxy growth is much more dependent on halo mass--high mass galaxies form early, growing slowly at low redshift, while low mass galaxies form late.  Slow late time growth of massive galaxies is necessary to reproduce the observed prevalence of quiescent galaxies as mass increases \citep[e.g.][]{vandenBosch:2008fv, Wetzel:2012lk}.  This is related to the downsizing phenomenon \citep{Neistein:2006fl}, wherein star-formaiton shifts to less massive galaxies at low redshifts \citep[e.g.][]{Kodama:2004gb, Jimenez:2005gj, Juneau:2005ig, Bell:2005hs, Bundy:2006em}.

It is the same story when comparing RV to any of the other mass based models.  In Fig. \ref{fig:mstar_zform}, we show the the median redshift galaxies formed a fraction, $f$, of their $z=0$ stellar mass (this is another way of examining the information in Fig. \ref{fig:mstar_of_z_rm}).  The top panel shows median redshift galaxies formed 10\% of their mass.  The middle and bottom panels show the redshift for 50\% and 90\% respectively.  The primary conclusion to draw from this is that the mass-based SHAM models generically predict much later growth compared to RV.  Interestingly, even the RM model, where the SMHM relation does not evolve, results in galaxy growth much closer to the evolving models than with RV.  This is a consequence of the fact that haloes form their potential wells early, primarily adding mass in the outskirts at late times \citep{Bosch:2014cu}.  As a result, any model for galaxy formation based on $V_{\rm peak}$ will have to decouple galaxy growth from the growth of $V_{\rm peak}$ more than models based on $M_{\rm peak}$.

@@

Given that the M13, Y12, and B13 models were fit to the stellar mass function at various redshifts and reproduce the cosmic star-formation density in the Universe (among other observables), it is difficult to imagine how to make a model like RV consistent with these same observational constraints.  While not explored here, it is likely that a more complicated $V_{\rm peak}$ based SHAM model could produce realistic galaxy growth histories, but this would require evolution in the SMVP relation.  Any evolution will require the SMVP relation be lower at high redshifts (as is the case for the SMHM relation in M13, Y12, and B13), reducing the stellar mass assigned to satellite galaxies in typical implementations.  This sort of modification would result in a decrease in the galaxy clustering signal on small scales.  We leave formulations of evolving $V_{\rm peak}$ to future work.

Given this failure of the $V_{\rm peak}$ SHAM, it is not clear that there is a reason to prefer a $V_{\rm peak}$ based SHAM model over models based on $M_{\rm peak}$.  Furthermore, apart from arguments related to fitting $w_p(r_p)$ and $\phi(M_*,z)$, there are other difficulties that must be overcome if $V_{\rm peak}$ is to be argued as more fundamental in driving the stellar mass mapping into (sub-)haloes.  For any given (sub-)halo, $V_{\rm peak}$ is generally set during major mergers \citep[1:5 or larger, ][]{Behroozi:2014bg}, and if stellar mass is tightly correlated with $V_{\rm peak}$, this could imply bursty star-formation closely tied to major mergers which is not favoured by observations \citep[e.g. star formation histories,][]{Diemer:2017gj}.  On the other-hand, $M_{\rm peak}$ remains a theoretically attractive as a quantity that should be tightly coupled to stellar mass.  Peak halo mass should be a good indicator of the amount of baryons that have been accreted and thus available to a galaxies to grow over its history.  Nevertheless, it is possible that feedback processes that result in the modulation of star-formation efficiency happen to correlate with $V_{\rm peak}$, but this remains unclear.

In our comparison between $V_{\rm peak}$-based SHAM and $M_{\rm peak}$-based SHAM it is clear that there is a tension between fitting galaxy clustering on small scales while simultaneously reproducing the build up of stellar mass in the Universe.  This tension presents a ``clustering crisis'' for SHAM.  Our exploration of the reasons for this tension suggests multiple ways SHAM models could be altered in order to alleviate this crisis.  We have identified too few satellites as the primary culprit for the clustering deficiency in mass based SHAM models.  One avenue to address this problem is to simply increase the number of satellites in the mass-based models as many models have found to be necessary.  In \S \ref{sec:orphans}, we examine the plausibility of missing sub-haloes, and thus satellites, in our implementation of the mass-based SHAM models.  A significant population of missing sub-haloes would be an indication that so called ``oprhan'' galaxies play an important role in solving the clustering crisis.  Apart from orphans, we also consider two other physically motivated modifications to increase the satellite fraction--post-accretion satellite growth in \S \ref{sec:sat_growth} and assembly bias effects in \S \ref{sec:assem_bias}.

@@

One possible solution to the lack of small scale clustering signal in SHAM models is to include a population of ``orphan'' galaxies.  Given the finite mass and force resolution of dark matter simulations, it is reasonable to expect that some sub-haloes are artificially disrupted or otherwise missing from the halo catalogues at z=0 \citep[][van den Bosch 2017 in prep.]{Carlberg:1994kl, vanKampen:1995ix, Guo:2013fm}.  Alternatively, sub-halo finders may fail in identifying sub-haloes when the density contrast is low \citep{Wetzel:2010ei, Muldrew:2011gr, Knebe:2011jc, Onions:2012iv, Knebe:2013bp, vandenBosch:2016ky} as is the case in the central regions of host haloes.  If this is in fact occurring, then it is appropriate to include a population of ``orphan'' galaxies, galaxies that have no identifiable sub-halo within a simulation but should rightfully be included if the simulation or sub-halo finder had been more successful.

After sub-haloes are accreted onto a more massive host-halo, mass is tidally stripped as the sub-halo orbits within the potential of its host, resulting in a ratio between the $z=0$ mass and the mass at accretion that is generally less than unity:
%
\begin{linenomath}
\begin{equation}
f_m = m_{\rm sub}/m_{\rm acc}
\end{equation}
\end{linenomath}
%
where the mass of a sub-halo, $m_{\rm sub}$, is the instantaneous mass that is bound to the sub-halo, and $m_{\rm acc}$ is the virial mass at the time of accretion.  Eventually, sub-haloes may simply be stripped below the mass resolution of the simulation, $f_m \times m_{\rm acc} \sim m_p$.  If it is common for sub-haloes to survive to this point, this will result in a need for orphans.  Alternatively, if sub-haloes are not tracked accurately, mass resolution is not sufficiently high, or the force resolution is not sufficient, some sub-haloes may become disrupted prematurely, meaning fewer sub-haloes will be available to host satellite galaxies when applying an abundance matching scheme.

@@

We calculate the sub-halo completeness in the simulation as the ratio of the empirical $f_{\rm sub}$ and the fit using eq. \ref{eq:fsub_fit} as shown in the bottom panel of Fig. \ref{fig:f_sub}.  We then model the completeness, $C_{\rm sub}$, as a function of halo mass as:
%
\begin{linenomath}
\begin{equation}
\label{eq:sub_completeness}
C_{\rm sub}(M_{\rm peak}) = \frac{C_0}{1.0 + \left(\frac{M_0}{M_{\rm peak}}\right)^{\gamma}}
\end{equation}
\end{linenomath}
%
We find $\log(M_0/h^{-1}M_{\odot})=9.980\pm0.003$ and $\gamma=1.27 \pm 0.02$ provides a good fit, as shown by the red line in the bottom panel in Fig. \ref{fig:f_sub}.  For now, we fix $C_0=1.0$, and we examine the possibility of $C_0<1.0$ at the end of this section.  The implicit assumption here is that when sub-haloes are well resolved, as is the case for massive (sub-haloes) at the time of infall, there should be no missing sub-structure.  Our estimation of the completeness is broadly consistent with the 1000 particle threshold found by \citet{Guo:2013fm} using the Millennium simulation suite (shown as the grey line in the bottom panel of Fig. \ref{fig:f_sub}).  Furthermore, we find that $C_{\rm sub}(M_{\rm peak})$ is very nearly constant with redshift (between $z=0-4$).

@@

Given the success of our modifications to the RM model to introduce assembly bias to $M_*$, we now consider modifications to the evolving mass-based modes.  As discussed, each of the evolving models makes the assumption that satellite galaxies at the time of accretion have the same mass as central galaxies in haloes with equal $M_{\rm peak}$.  In these models, correlating $M_*$ with $z_{\rm f}$ will increase the mass of satellite galaxies if haloes which become sub-haloes form significantly earlier than haloes which remain host-haloes.  This differs from the model for post-accretion satellite growth discussed in \S \ref{sec:sat_growth}.  In this model, galaxies which become satellites grow more quickly than galaxies which remain centrals.  Therefore, satellites are ``over-massive'' compared to centrals at $z_{\rm acc}$.
